Reproducible methods for modeling combined public transport and cycling trips and associated benefits: \ evidence from the biclaR tool

### 2024-05-08 paper/PaperCEUS/PaperCEUS_rev1_word.Rmd
final rev1 CEUS paper
submitted! closes #47
closes #48
closes #49
closes #50

---
title: "Reproducible methods for modeling combined public transport and cycling trips and associated benefits: \ evidence from the biclaR tool"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa. Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using Health World Organization tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of all trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    # citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: black
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---
\hypersetup{citecolor=black}

@@

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA.
After presenting the [Methods] used in Chapter 2, we assess its socio-environmental impacts using open data and open-source tools, presenting the [Results and Discussion] in Chapter 3, followed by Chapter 4 where we discuss the [Generalizability and Limitations] of the methods.
We [Conclude](#Conclusions) the paper in Chapter 5 with the main findings and policy implications.

>>>>

---
title: "Reproducible methods for modeling combined public transport and cycling trips and associated benefits: \ evidence from the biclaR tool"
author:
  - name: Rosa Félix\corref{cor1}
    affiliation: CERIS
    correspondingauthor: true
    # footnote: 1
    email: rosamfelix@tecnico.ulisboa.pt
  - name: Filipe Moura
    affiliation: CERIS
    email: fmoura@tecnico.ulisboa.pt
  - name: Robin Lovelace
    affiliation: ITS
    email: r.lovelace@uleeds.uk
address:
  - code: CERIS
    address: "CERIS, Instituto Superior Técnico - Universidade de Lisboa. Av Rovisco Pais 1, 1049-001 Lisboa, Portugal"
  - code: ITS
    address: "Institute for Transport Studies, University of Leeds. 34-40 University Rd, Leeds LS2 9JT, UK"
abstract: |
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using Health World Organization tools. The impacts of shifting car trips to public transport were also estimated and monetized.
  The results show that 10% of all trips could be made by cycling in combination with public transport. Shifting to cycling for the shorter first and last mile stages can reduce annual CO~2~eq emissions from 3,000 to 7,500 tons/day, while for the public transport leg, the transfer from car avoids of up to 20,500 tons of CO~2~eq emissions per year. The estimated socio-environmental benefits are of €125 million to €325 million over 10 years.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
  \
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
journal: "Computers, Environment and Urban Studies"
date: "`r Sys.Date()`"
classoption: review, doubleblind, 3p, authoryear #change later to preprint instead of review
bibliography: bibliography.bib
biblio-style: elsarticle-harv
# linenumbers: true
numbersections: true
# Use a CSL with `citation_package = "default"`
csl: https://www.zotero.org/styles/computers-environment-and-urban-systems
# csl: computers-environment-and-urban-systems.csl
preamble: |
  \biboptions{authoryear}
  \usepackage{url}
  \usepackage{float}
  \usepackage{booktabs}
  \usepackage{longtable}
  \usepackage{makecell}
  \usepackage{multirow}
output: 
  rticles::elsevier_article:
    keep_tex: yes
    # citation_package: "default" #ou nenhum
    latex_engine: xelatex
    extra_dependencies: "subfig"
#   bookdown::word_document2:
#     fig_caption: true
# always_allow_html: true
urlcolor: blue
citecolor: black
linkcolor: blue
# natbiboptions: longnamesfirst,angle,semicolon
---
\hypersetup{citecolor=black}

@@

This paper estimates the potential for combining cycling and PT to substitute car trips in the LMA.
After presenting the [Methods] used in Section 2, we assess its socio-environmental impacts using open data and open-source tools, presenting the [Results and Discussion] in Section 3, followed by Section 4 where we discuss the [Generalizability and Limitations] of the methods.
We [Conclude](#Conclusions) the paper in Section 5 with the main findings and policy implications.
